---
title: "TP4"
format: html
editor: visual
---

## Grade analysis

```{r, message=FALSE, warning=FALSE, echo=FALSE}
library(here)
library(dplyr)
library(tidyr)
```

### 1)

```{r, message=FALSE, warning=FALSE, echo=FALSE}
here::i_am('r-101-grade-analysis.Rproj')
grades<-here('grades.csv')
grades<-read.csv(grades)

```

### 2)

```{r, message=FALSE, warning=FALSE, echo=FALSE}
t<-grades|>select(Exam)|>summarize(max=max(Exam, na.rm = T), med=median(Exam, na.rm = T), min=min(Exam,na.rm=T))


knitr::kable(t)


```

### 3)

```{r, message=FALSE, warning=FALSE, echo=FALSE}

absent<-grades|>select(Exam)|>
  summarise(Nombre_NA = sum(is.na(Exam)))
  


```

```         
Il y avait `r absent` personnes absentes à l'examen
```

### 4)

```{r, message=FALSE, warning=FALSE, echo=FALSE}

library(ggplot2)


exam_noNA<-grades|>select(Exam)|>na.omit()

ggplot(exam_noNA, aes(x=Exam))+
  geom_histogram(bin=30)+
  xlab("grade")

```

### 5)

```{r, message=FALSE, warning=FALSE, echo=FALSE}
table<-grades %>% group_by(Group) %>% summarise(count=n())

knitr::kable(table)
```

### 6)

```{r, message=FALSE, warning=FALSE, echo=FALSE}



ggplot(grades, aes(x=Group))+
  geom_bar(bin=30)+
  xlab("group")+
  theme(axis.text.x=element_text(angle = 45,hjust=1))
```

### 7)

```{r, message=FALSE, warning=FALSE, echo=FALSE}

ggplot(grades, aes(x=Exam))+
  geom_histogram(bin=30)+
  xlab("group")+
  facet_wrap(~Group)
  

```

```{r, message=FALSE, warning=FALSE, echo=FALSE}
ggplot(grades, aes(x = Group, y = Exam, fill=Group)) +
  geom_boxplot() +
  xlab("Groupe") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) 
 
```

### 8)

```{r, message=FALSE, warning=FALSE, echo=FALSE}

missed_exam <- grades %>% filter(is.na(Exam))

missed_exam_count<- missed_exam %>%
  group_by(Group) %>%
  summarise(Num_Missed_Exam = n())

knitr::kable(missed_exam_count)
```

### 9)

```{r, message=FALSE, warning=FALSE, echo=FALSE}
ggplot(missed_exam_count, aes(x = Group, y = Num_Missed_Exam)) +
  geom_col(stat = "identity", fill = "blue") +
  xlab("Groupe") +
  ylab("Nombre d'Absents") +
  theme_minimal()+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

### 10)

```{r, message=FALSE, warning=FALSE, echo=FALSE}
grades_long <- pivot_longer(grades, 
                            cols = -c(Id, Group), #Colonnes d'identifiants (non pivotées)
                            names_to = "name",  # Nouvelle colonne pour stocker les noms des colonnes pivotées
                            values_to = "Values")# Nouvelle colonne pour stocker les valeurs pivotées


knitr::kable(grades_long)
```

### 11)

```{r, message=FALSE, warning=FALSE, echo=FALSE}
missed_exam_stud <- grades_long %>% filter(is.na(Values))

missed_exam_count_stud<- missed_exam_stud %>%
  group_by(Id) %>%
  summarise(Num_Missed_Exam_stud = n())

knitr::kable(missed_exam_count_stud)

```

The students that missed no exam do not appear.

### 12)

```{r, message=FALSE, warning=FALSE, echo=FALSE}
ggplot(missed_exam_count_stud, aes(x = Id, y = Num_Missed_Exam_stud)) +
  geom_bar(stat = "identity", fill = "orange") +
  xlab("student ID") +
  ylab("Number of missed exams") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

### 13)

```{r, message=FALSE, warning=FALSE, echo=FALSE}
missed_exam_long <- grades_long %>% filter(is.na(Values))

missed_exam_count_long<- missed_exam %>%
  group_by(Group) %>%
  summarise(Num_Missed_Exam = n())

knitr::kable(missed_exam_count_long)
```

### 14)

```{r, message=FALSE, warning=FALSE, echo=FALSE}

library(stringr)

```

```{r, message=FALSE, warning=FALSE, echo=FALSE}
missed_online_stud<-grades_long|>filter(str_starts(name,'Online'))|> filter(is.na(Values))

missed_online_count_stud<- missed_online_stud %>%
  group_by(Id) %>%
  summarise(Num_Missed_online_stud = n())

knitr::kable(missed_online_count_stud)


```

We can spot that the students are more likely to miss a online exam.

### 15)

```{r, message=FALSE, warning=FALSE, echo=FALSE}
exam_stud<-grades_long|>filter(name=="Exam") |>select(Id,Values)|>arrange(Id)#df avec les notes par étudiant triés par id croissant comme à la question précédente.

df<-merge(exam_stud,missed_online_count_stud, by="Id")

knitr::kable(df)

```

We now have a dataframe df counting the number of missed online exam and the final exam grade per student.

Let's do the representation :

```{r, message=FALSE, warning=FALSE, echo=FALSE}

ggplot(df, aes(x = Num_Missed_online_stud, y = Values, color=Values)) +
  geom_point()+
  xlab("Number of missed online exam")+
  ylab("Exam grade")
```

In this graph, students are grouped by the number of online exam they have missed.

We can clearly spot that the more online exams are missed, the more the distribution of final exam grades is concentrated between 0 and 10.

### 16)

Starting from the original data set, we add a column Missed (True if there is a NA in the line for each student).

We take into account every grade including the online ones.

Then we order the new data frame by Id.

```{r, message=FALSE, warning=FALSE, echo=FALSE}
df2<-grades %>%
  mutate(Missed = rowSums(is.na(.)) > 0)|>group_by(Id)|>arrange(Id)

#df2

table<-df2|>select(Id,Missed)

knitr::kable(table)

```

### 17)

```{r, message=FALSE, warning=FALSE, echo=FALSE}

df_P_missed<-df2|>select(Id, Group, Missed)|>group_by(Group)|>summarize(P_missed = mean(Missed == TRUE) * 100)

knitr::kable(df_P_missed)
```

### 18)

```{r, message=FALSE, warning=FALSE, echo=FALSE}


#We create a df with the average exam grade per group.

df_mean<-df2|>select(Id, Group, Exam)|>group_by(Group)|>summarize(mean_exam = mean(Exam, na.rm = TRUE))


```

```{r, message=FALSE, warning=FALSE, echo=FALSE}

#we create a df with the two variable of interest :

df3<-cbind(df_mean,df_P_missed[2])

#We do the plot


#df3

ggplot(df3, aes(x = P_missed, y = mean_exam, color=mean_exam)) +
  geom_point()+
  xlab("P_missed")+
  ylab("mean exam group")


```

Once again, we can spot that the higer pourcentage of people having missed at least one exam is, the more the distribution of the mean per group is concentrated between 0 and 8.

We could even pretend to see a negative linear correlation between these two variables.
